### 2022-06-13 Manuscript/Manuscript.Rmd
Visualize simulation results

---
title: "Non-convergence in iterative imputation"
author: "H.I. Oberman"
output: pdf_document
bibliography: references.bib  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

# filters
coeff <- "X4"
maxit <- 20
```

```{r data}
performance <- readRDS("./Results/performance.RDS")
summary <- performance %>% 
  group_by(method, mech, prop, .it, term) %>% 
  summarise(across(c(bias, cov, ciw, ac, psrf), mean, na.rm = TRUE))
SEs <- performance %>% 
  group_by(method, mech, prop, .it, term) %>% 
  summarise(across(c(bias, ciw, ac, psrf), var, na.rm = TRUE)) %>% 
  mutate(across(c(bias, ciw, ac, psrf), sqrt))
results <- left_join(summary, SEs, by = c("method", "mech", "prop", ".it", "term"), suffix = c("", "_SE")) %>% 
  mutate(cov = cov,
         cov_SE = sqrt((cov * (1 - cov)) / 2000))
```

# Abstract

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

# Introduction

- Theoretical background

- Research question

# Methods

- Simulation design (ADEMP)

- Simulation checklist (Table \ref{table:check})

Pseudo-code:

```
for (data geneneration conditions) {
  generate complete data
  for (amputation conditions) {
    ampute complete data
    for (imputation conditions) {
      impute incomplete data
      analyze imputed data
      }}}
```

\begin{table}[ht]
\caption{Checklist for reporting on imputation methodology evaluations.}
\label{table:check}
\begin{tabular}{ll}
% \hline
%   & Suggested aspects to report                                                       \\
\hline
1 & Simulation scope                                                                   \\ \cline{2-2}
  & Aim                                                                                \\
  & Design (incl. pseudo-code or flow diagram)                                         \\
  & Number of simulation repetitions                                                   \\ \hline
2 & Comparative truth                                                                  \\ \cline{2-2}
  & Data-generating mechanism (model-based or design-based)                            \\
  & Sampling variance                                                                  \\
%  & Data characteristics (number of observations, number of variables, variable types, \\
%  & coherence between variables incl. data structures e.g. clustering)                 \\
  & Data characteristics (incl. multivariate relations and structures e.g. clustering) \\
  & Estimand                                                                           \\ \hline 
3 & Induced missingness                                                                \\ \cline{2-2}
  & Missingness mechanism (incl. type or functional form of the missing data   model)  \\
  & Missingness pattern (incl. missingness proportion)                                 \\ \hline 
4 & Applied methods                                                                    \\ \cline{2-2}
  & Imputation methods (incl. parameters e.g. the number of imputations)               \\
  & Analytic methods (incl. calculation of standard errors e.g. pooling   rules)       \\
  & Reference method (e.g. complete case analysis)                                    \\ \hline 
5 & Imputation evaluation                                                              \\ \cline{2-2}
  & Imputation-generating process (e.g. algorithmic non-convergence)                   \\
  & Imputation model fit (e.g. posterior predictive checks)                            \\
  & Distributional characteristics (e.g. plausibility of imputed values)               \\ \hline 
6 & Performance evaluation                                                             \\ \cline{2-2}
  & Statistical properties (e.g. confidence validity)                                  \\
  & Comparative performance (e.g. predictive accuracy)                                 \\ \hline
\end{tabular}
\end{table}

# Results

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

```{r}
# results %>% 
#   filter(term == "X2", .it < 26) %>% 
#   ggplot(aes(x = .it, y = bias)) +
#   geom_hline(yintercept = 0, color = "grey", size = 2) +
#   geom_line(color = "blue") +
#   geom_line(aes(x = .it, y = bias + bias_SE), color = "lightblue", linetype = "dashed") +
#   geom_line(aes(x = .it, y = bias - bias_SE), color = "lightblue", linetype = "dashed") +
#   facet_grid(as.factor(prop) ~ mech) +
#   theme_classic()

results %>% 
  filter(term == coeff, .it > 0, .it < maxit + 1) %>% 
  ggplot(aes(x = .it, y = bias)) +
  geom_hline(yintercept = 0, color = "black") +
  geom_errorbar(aes(ymin = bias - bias_SE, ymax = bias + bias_SE), alpha = 0.1) +
  geom_line(color = "blue") +
  geom_point(color = "blue") +
  facet_grid(as.factor(prop) ~ mech) +
  theme_classic()

results %>% 
  filter(term == coeff, .it > 0, .it < maxit + 1) %>% 
  ggplot(aes(x = .it, y = ac)) +
  geom_hline(yintercept = 0, color = "black") +
  geom_errorbar(aes(ymin = ac - ac_SE, ymax = ac + ac_SE), alpha = 0.1) +
  geom_line(color = "blue") +
  geom_point(color = "blue") +
  facet_grid(as.factor(prop) ~ mech) +
  theme_classic()

results %>% 
  filter(term == coeff, .it > 0, .it < maxit + 1) %>% 
  ggplot(aes(x = .it, y = psrf)) +
  geom_hline(yintercept = 1, color = "black") +
  geom_errorbar(aes(ymin = psrf - psrf_SE, ymax = psrf + psrf_SE), alpha = 0.1) +
  geom_line(color = "blue") +
  geom_point(color = "blue") +
  facet_grid(as.factor(prop) ~ mech) +
  theme_classic()
```

# Discussion

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

# References